\section{Time as a Derived Quantity}

\subsection{Problem Class}

Time is commonly treated as a fundamental parameter.
Within the MACHWERK framework, time is not fundamental.

Time is a derived, projected ordering relation.

The framework does not ask what time \emph{is},
but under which conditions time statements are admissible.

\subsection{Absence of Time in the Relational Space}

In the relational full space $\mathcal{U}$:

\begin{itemize}
\item there is no global time parameter
\item there is no privileged ordering
\item there is no universal clock
\end{itemize}

$\mathcal{U}$ contains relations, not sequences.

Temporal structure does not exist prior to projection.

\subsection{Emergence via Projection}

Let
\[
\Pi_t : \mathcal{U} \rightarrow m_2
\]
be a projection into an observable calculation domain
where ordered process comparisons are defined.

Time arises through ordered comparison of process rates:
\[
t \sim \frac{R_i}{R_j}
\]
or through cyclic reference processes.

Time is therefore not measured directly,
but inferred relationally.

\subsection{Local Validity of Temporal Ordering}

Temporal ordering is only valid within domains
where the projection $\Pi_t$ is injective.

Distinct relational configurations may project
to identical temporal orderings:
\[
\Pi_t(u_1) = \Pi_t(u_2)
\quad \text{with} \quad u_1 \neq u_2
\]

Thus, time does not uniquely encode relational structure.

\subsection{Black Boundary for Time}

Define
\[
\Sigma_t := \{ u \in \mathcal{U} \mid \Pi_t \text{ is non-injective} \}
\]

Beyond $\Sigma_t$:

\begin{itemize}
\item temporal ordering remains computable
\item durations remain definable
\item causal reconstruction fails
\end{itemize}

Time continues to function numerically,
but loses interpretive exclusivity.

\subsection{CRA Constraint}

A temporal statement $S(t)$ is admissible
only if it depends solely on projected relations
and is invariant under relational substitutions
that preserve $\Pi_t$.

Absolute or global time claims violate CRA.

\subsection{Formal Conclusion}

Time in MACHWERK is not a background structure.

It is a projection-stable ordering relation
derived from comparative process behavior.

Time exists where projection is injective,
and becomes formally ambiguous beyond the black boundary.
