\section{Energy as a Relational Rate Measure}

\subsection{Problem Class}

In standard physics,
energy is treated as a conserved scalar quantity
associated with systems or fields.

Within MACHWERK,
energy is not fundamental.

Energy is a relational rate measure
arising from projection.

\subsection{Absence of Energy in $\mathcal{U}$}

The relational full space $\mathcal{U}$ contains:

\begin{itemize}
\item no energy values
\item no conserved scalars
\item no absolute resource quantities
\end{itemize}

Relations in $\mathcal{U}$ encode
comparative process structure only.

\subsection{Emergence via Projection}

Let
\[
\Pi_E : \mathcal{U} \rightarrow m_2
\]
be a projection into a domain
where process changes are temporally ordered.

Energy appears as a comparative rate of change:
\[
E \sim \frac{\Delta R}{\Delta \lambda}
\]

Energy measures how rapidly
relational configurations change
relative to an ordering parameter.

\subsection{Energy as Bookkeeping Quantity}

Energy is not causal.
It is a bookkeeping invariant
within a stable projection.

Conservation of energy reflects
projection consistency,
not a fundamental law of nature.

\subsection{Context Dependence}

Different projections
yield different energy assignments
for the same relational configuration.

This does not imply contradiction,
only domain dependence.

Energy is locally meaningful,
not globally absolute.

\subsection{Black Boundary for Energy}

Define
\[
\Sigma_E := \{ u \in \mathcal{U} \mid \Pi_E \text{ is non-injective} \}
\]

Beyond $\Sigma_E$:

\begin{itemize}
\item rate relations remain formalisable
\item multiple configurations map to identical energies
\item conservation loses physical meaning
\end{itemize}

Formal calculation persists.
Physical interpretation does not.

\subsection{CRA Constraint}

An energy-related statement $S(E)$
is admissible only if it depends solely
on projected relational rates
and remains invariant under
non-injective substitutions.

Intrinsic energy assumptions
violate CRA.

\subsection{Formal Conclusion}

Energy in MACHWERK is not a substance
and not a universal invariant.

It is a projection-dependent
rate descriptor within $m_2$.

Beyond its black boundary,
energy remains calculable
but ceases to carry physical significance.
